FINE SCALE REGISTRATION OF SIDEWALKS AND OTHER RIBBON-LIKE FEATURES FROM AERIAL IMAGERY

### 2018-11-06 Chapters/2.Theory/Theory.tex
Corrections and updates to notation

A variety of approaches are able to locate precise object boundaries, although few directly address the problem of sidewalk segmentation in "blobby" or compact regions. 
Several tools for segmentation appear for segmenting ribbon-like features in \ac{VHR}, we list Grabcut ~\cite{Rother2004-ou}, Slic~\cite{Achanta:149300}, and Active Contours~\cite{Kass88snakes:active} as examples.
Results for these methods on a simple sidewalk are shown in \figref{fig:Method_comparison} which shows qualitative results of each method. 
We observe that grabcut had decent performance on this example, but it is still unable to identify edges in shadow or correctly determine the width of the sidewalk region in the vicinity of a driveway, which seems to be made of the same material as the sidewalk. 
In addition we observe that all three approaches recognize the gutter as part of the sidewalk because it has the camouflage texture.

\begin{figure}[H]
    \centering
    \includegraphics[width=0.55\textwidth]{Figures/compare.png}
    \caption[Method comparison with Grabcut, Active Contours, and Slic]{
        % Method comparison with Grabcut, Active contour and Slic
        A comparison between Grabcut(left), SLIC (center) and Active Contours (right) 
        on an example of a sidewalk from 6-inch aerial imagery; 
        \aclp{TP} are rendered in green, 
        \aclp{FN} are rendered in blue, and 
        \aclp{FP} are rendered in red.
        The SLIC image contains all super-pixels that intersect the initial trajectory of a walking path; the Active Contours and GrabCut methods are initialized with the SLIC results.
    }
    \label{fig:Method_comparison}
\end{figure}

@@

\begin{figure}[H]
    \centering
    \includegraphics[width=\textwidth]{Figures/grabcut_sample.png}
    \caption[Example of Grabcut]{Detail demonstration with grabcut method. Row 2 indicates examples of false positive, which is recognized a non-sidewalk feature (shadow) and similar texture (driveway) as a part of sidewalk.}
    \label{fig:grabcut}
\end{figure}

Grabcut is a powerful image segmentation tool that combines both texture(color) and edge(contrast) information~\cite{Rother2004-ou}. 
It was designed to efficiently extract a foreground object from a complex background. 
According to figure \ref{fig:Method_comparison} and figure \ref{fig:grabcut}, the overall performance on grabcut applied to a ribbon-like feature is fairly accurate, but it's hard to identify objects above the sidewalk such as shadows, gutters or foliage. 
Grabcut seems to identify the occluded regions as negative, we suspect that is because of the lack of width control.

@@

\begin{figure}[H]
    \centering
    \includegraphics[width=\textwidth]{Figures/slic_sample1.pdf}
    \caption[Example of Simple Linear Iterative Clustering]{Detail demonstration on segmenting large image into irregularly shaped super-pixels with \ac{SLIC}. Row 2 indicates examples of false positive, with recognized a non-sidewalk feature (shadow) as a part of sidewalk.}
    \label{fig:slic}
\end{figure}

Simple linear iterative clustering (SLIC) works with super-pixels, which segment a large image into different irregularly shaped regions of homogeneous appearance or texture~\cite{Achanta:149300}. 
It uses the compact evenly distributed homogeneous regions of an image as a pre-processing step, so that image segmentation can be accomplished by grouping super-pixels which snap to natural boundaries in an image. 
Alternatively, super-pixels can initiate a region growing process~\cite{Borovec2017-fz}. 
As shown in figure \ref{fig:slic}, slic will identify a non-sidewalk feature as positive if there is a shadow covering the sidewalk, which may lead to over mark the boundary (\aclp{FP}).
This becomes a problem since all super-pixels that we decide to use are based on a line in between both edges of sidewalk which we marked as ground truth. 
On the one hand, if the super-pixel above lay on the line, we got an error, since we marked part of the feature that not the sidewalk on the sidewalk, which gives us \aclp{FP}. 
On the other hand, if that super-pixel does not lay on the line, we missed a part that's clear is the sidewalk. Thus, finding a way to separate such a feature become important. 
Trying to decrease the size of the width of each super-pixel was one of the ways to fix it but may cause an incompleteness since each super-pixel is smaller than the width of the sidewalk.
So the idea of width control become necessary.

@@

\begin{figure}[H]
    \centering
    \includegraphics[width=\textwidth]{Figures/ac1.png}
    \caption[Example of Active Contours]{Detail demonstration with active contours method. Row 2 indicates examples of false positive, which is recognized a non-sidewalk feature (shadow) and similar texture (driveway) as a part of sidewalk.}
    \label{fig:ac}
\end{figure}

A very successful region growing process is snakes, or Active Contours, which starts with an approximate contour and evolves the boundary to optimize an energy function that favors likely contour shapes. 
Compared with the other two segmentation tools, active contour may not be the best solution for identifying ribbon-like feature since it is not the best tool to separate blurred edges.
Al-Diri~\cite{ActiveContou09} proposed using active contour to identify each edge separately, in the mean time keep maintaining width consistency. 
It is a good approach but it is not ideal to apply on our problem, since their ribbon-like feature (vessels) does not requiring predict edges under obstacles, such as trees or cars that above the sidewalk.

@@

Another image segmentation approach that has seen extraordinary success is the use of Convolutional Neural Networks (CNNs) and transposed convolution~\cite{Badrinarayanan2017-il, Noh2015-ni, Shelhamer2017-rf, Ronneberger2015-sv}. 
Unlike the proposed approach, these have millions of tunable parameters and need significant pre-annotated data and offline training to build a model. 
In a recent exploration of learning to segment street networks from online maps~\cite{Kaiser2017-np}, built a fully convolutional network with 137 million parameters.
A similar approach with another deep architecture called U-net~\cite{Ronneberger2015-sv} was applied to the street network segmentation problem~\cite{Zhang2017-gi}.
These approaches claim to be robust to occlusion and ambiguous regions, but they benefit from a large amount of carefully annotated data that includes road width~\cite{Mnih2013-dp}. 
This type of pre-annotated data with aligned aerial imagery is much harder to obtain for walking paths. 
\ac{DPST} has only handful of tunable parameters with clear interpretations, so that they can be interactively set and used without offline training.

>>>>

We now turn our attention to competing prior art that would seem to address the same or a similar problem to the proposed approach. 
A variety of approaches are able to locate precise object boundaries, although few directly address
the problem of sidewalk segmentation. Several tools for segmentation
appear for segmenting features in \ac{VHR}, we list Grabcut ~\cite{Rother2004-ou},
SLIC~\cite{Achanta:149300}, and Active Contours~\cite{Kass88snakes:active} as examples. 
Results for
these methods on a simple sidewalk are shown in \figref{fig:Method_comparison} which shows
qualitative results of each method. We observe that \GrabCut{} had decent performance on this example,
but it is still unable to identify edges in shadow or correctly determine the width of the sidewalk
region in the vicinity of a driveway, which seems to be made of the same material as the sidewalk.
In addition we observe that all three approaches recognize the gutter as part of the sidewalk
because it has a similar texture.

\begin{figure}[H]
    \centering
    \includegraphics[width=0.55\textwidth]{Figures/compare.png}
    \caption[Method comparison with Grabcut, Active Contours, and Slic]{
        % Method comparison with Grabcut, Active contour and Slic
        A comparison between \GrabCut{}~(a), SLIC~(b) and \ActiveContours{}~(c) 
        on an example of a sidewalk from 6-inch aerial imagery; 
        \aclp{TP} are rendered in green, 
        \aclp{FN} are rendered in blue, and 
        \aclp{FP} are rendered in red.
        The SLIC image contains all super-pixels that intersect 
        the initial trajectory of a walking path; 
        the \ActiveContours{} and \GrabCut{} methods are initialized with the SLIC results.
    }
    \label{fig:Method_comparison}
\end{figure}

@@

\begin{figure}[H]
\centering 
\includegraphics[width=\textwidth]{Figures/grabcut_sample.png} 
\caption[Example of \GrabCut{}]{
Details of our demonstration of the \GrabCut{} method. Row 2 indicates examples of errors, which are
\ac{FN} pixels recognized a non-sidewalk feature (shadow) and \ac{FP} pixels that have a similar
texture (driveway) as a part of sidewalk.}
\label{fig:grabcut}
\end{figure}

\GrabCut{} is a powerful image segmentation tool that combines both texture(color) and edge(contrast) information~\cite{Rother2004-ou}. 
It was designed to efficiently extract a foreground object from a complex background. 
As we illustrate in \figurename{} \ref{fig:Method_comparison} and \figurename{} \ref{fig:grabcut}, the overall
 performance on \GrabCut{} applied to a ribbon-like feature is fairly accurate, but it tends to to fail at identifying objects near the 
 sidewalk such as shadows, gutters or foliage. 
\GrabCut{} seems to identify the occluded regions as negative, we suspect that is because of the lack of width control.

@@

\begin{figure}[H]
\centering
\includegraphics[width=\textwidth]{Figures/slic_sample1.pdf}
\caption[Example of Simple Linear Iterative Clustering]{
Detail of our demonstration of segmenting large image into irregularly shaped super-pixels with \ac{SLIC}.
Row 2 indicates examples of errors, with a non-sidewalk feature (shadow) recognized as a part of the sidewalk.}
\label{fig:slic}
\end{figure}

\acf{SLIC} is a method for identifying super-pixels, which segment a large image into different
irregularly shaped regions of homogeneous appearance or texture~\cite{Achanta:149300}. It finds
compact evenly distributed homogeneous regions of an image as a pre-processing step, so that image
segmentation can be accomplished by grouping super-pixels which snap to natural boundaries in an
image. Alternatively, super-pixels can initiate a region growing process~\cite{Borovec2017-fz}. As
shown in figure \ref{fig:slic}, \ac{SLIC} will identify a non-sidewalk feature as positive if there
is a shadow covering the sidewalk, which may lead to over mark the boundary (\acp{FP}). 
The results we show \ac{SLIC} segmentation are superpixels that cross the original \ac{GIS} line strip.
When strong edges exist within the sidewalk (e.g. from shadows or occlusion) the \ac{SLIC} approach may
 fail to recognize part of the sidewalk (if the superpixel does not cross the line strip) or it may 
 detect objects that extend outside the sidewalk (if the superpixel crosses the line strip). 
Thus, finding a way to separate such a feature become
important. Trying to decrease the size or the width of each super-pixel is one of the ways to fix
it but may cause an incompleteness since each super-pixel is smaller than the width of the sidewalk.
So the idea of width control become necessary.

@@

\begin{figure}[H]
\centering
\includegraphics[width=\textwidth]{Figures/ac1.png}
\caption[Example of Active Contours]{
Detail demonstration of the \ActiveContours{} method. 
Row 2 indicates examples of false positive, which is recognized a non-sidewalk feature 
(shadow) and similar texture (driveway) as a part of sidewalk.}
\label{fig:ac}
\end{figure}

A very successful region growing process is snakes, or Active Contours, which starts with an approximate contour and evolves the boundary to optimize an energy function that favors likely contour shapes. 
%Compared with the other two segmentation tools, active contour may not be the best solution for identifying ribbon-like feature since it is not the best tool to separate blurred edges.
In general, it is biased towards segmenting compact regions with clear boundaries; but it has been used to segment blood vessels (which are ribbon-like features). 
Al-Diri~\cite{ActiveContou09} proposed using active contour to identify each edge of a vessel separately, while maintaining width consistency. 
The approach reported favorable results on blood vessel segmentation, but it is not ideal to apply on our problem, 
since their ribbon-like feature (vessels) does not requiring predict edges under obstacles,
 such as trees or cars that occlude the sidewalk.

@@

Another image segmentation approach that has seen extraordinary success is the use of Convolutional Neural Networks (CNNs) and transposed convolution~\cite{Badrinarayanan2017-il, Noh2015-ni, Shelhamer2017-rf, Ronneberger2015-sv}. 
Unlike the proposed approach, these have millions of tunable parameters and need significant pre-annotated data and offline training to build a model. 
In a recent exploration of learning to segment street networks from online maps~\cite{Kaiser2017-np}, built a fully convolutional network with 137 million parameters.
A similar approach with another deep architecture called U-net~\cite{Ronneberger2015-sv} was applied to the street network segmentation problem~\cite{Zhang2017-gi}.
These approaches claim to be robust to occlusion and ambiguous regions, but they benefit from a large amount of carefully annotated data that includes road width~\cite{Mnih2013-dp}. 
This type of pre-annotated data with aligned aerial imagery is much harder to obtain for walking paths. 
\ac{DPRST} has only handful of tunable parameters with clear interpretations, so that they can be interactively set and used without offline training.

### 2018-11-06 Chapters/3.Hypothesis/Hypothesis.tex
Corrections and updates to notation

We import LiDAR data and shape-file that contain geometric information of the location for each sidewalk. 
Instead of analysis the texture difference between desire sidewalk and non-sidewalk feature, we treat it as the ribbon-like feature and try to segment them base on their growth from start point.

@@

In order to simplify the problem, we first create a straightened "ribbon" image in which the sidewalk runs in a straight line, even if the original sidewalk if curved.
We represent a ribbon of length $L$ using parameters along and orthogonal to a medial curve $\mathbf{x}(s)$, where $s \in [0, L]$. The ribbon image is the set of all points within a perpendicular distance $r(s)$ to the medial curve.

@@

\begin{figure}[H]
    \centering
    \includegraphics[width=\textwidth]{Figures/sw-figure.pdf}
    \caption[2D Ribbon Image]{A demonstration of the parameters for a ribbon in 2D. $k_n$ indicates a parameter that describes the radius of the medial curve that we considered to be part of the ribbon, $k_f$ is a distance from the medial curve that we considered very unlikely to be part of the ribbon, and $k_W$ is the maximum width of the ribbon image.}
    \label{fig:2d_ribbon}
\end{figure}

As shown in figure \ref{fig:2d_ribbon}, from figure \ref{fig:osm_oxford_path}, we should assume that there is some uncertainty in the position $\mathbf{x}_s$ of the medial curve and also uncertainty in the radius $r_s$ of the discredited ribbon. 
But that there is range of distances $t \in 0\dots k_n$ from the medial curve which we can accept labeling as positive examples that are within the ribbon, and that there is some range of distances $t \in k_f\dots k_W$ which we can accept labeling as negative sample. 
Finally a range $t\in (k_n+1)...(k_f-1)$ of distances that are possible locations of the ribbon's boundary.
Therefore, when searching for improved medial curves and radii of the ribbon, we can consider a discrete range of offsets $t=-k_W,..., k_W$ chosen to be wide enough that true ribbon image itself likely fits entirely within the grid of input negative samples along its edges.

The orthorectified color image $\image(x,y)$
records the color at the point $x,y$ in the image plane.
Then $\image(\mathbf{x}_{s,t}),$ is a warped version of the image in which the ribbon is centered on a single strip at $t=0$.
The situation is illustrated in figure  \ref{fig:ribbon_3d}.  
In this warped image, we can estimate the amount $t_s$ to shift the ribbon's center in the $t-$axis, which is orthogonal to the curve,  and also the amount the radius $r_s$ so that the  ribbon at each row is in columns $(t_s - r_s)\dots(t_s+r_s)$.  
Our problem is to estimate the most likely values of $t_s$ and $r_s$.

@@

We aim to find a segmentation that combines \textit{shape} priors on the curvature and thickness of the ribbon-like features and also maximizes the likelihood of observing colors in the image. 
Let $\Lambda$ be a random variable for the event that a color $\lambda$ is observed in an image, and let $R$ indicate a random variable so that $R=1$ if a pixel is part of ribbon and $R=0$ if a pixel is not. 
A variety of approaches exist to learn the conditional probability $\Pr(\Lambda|R)$ of observing a color conditioned on whether a pixel within or outside of a ribbon \cite{hartigan1979algorithm, jordan1999introduction}.
Many of these methods take as input a set of training examples in order to choose parameters for some model.
Our conjecture is that the initial estimate of our sidewalk's mid-line, while perhaps not perfect, is much more likely to cover pixels that are part of a sidewalk than the pixels at the edge of the ribbon image.

We use this assumption to estimate the probability that a pixel is part of a ribbon conditioned on its color.
Let $k_n$ be a parameter that describes the radius around the medial curve that we will consider to be part of the ribbon with very high confidence. 
Let $k_f$ indicate a distance from the medial curve that we consider very unlikely to be part of the ribbon. 
Finally, let $k_W$ indicate maximum value of $t$ to consider when learning the distribution of colors in a ribbon image. 
We model the distribution of colors within the sidewalk as a Gaussian Mixture Model of $k_g=4$ full-co-variance gaussian distribution and we model the distribution of colors outside of the ribbon as a mixture $2k_g$ Gaussian since we expect more variation and complexity in this zone. 
As shown in figure \ref{fig:GMM_result}, given an initial estimate for the set of colors inside or outside of the ribbon, we estimate the weights and co-variances of the Gaussians using the \ac{EM-GMM} \cite{sridharan2014gaussian} algorithm.

@@

We will assume that widths may take on values from a set of predefined odd numbers (so that there is always a pixel in the center) 
Then the negative log-likelihood of for a particular pair of values $t_s$ and $r_s$ at a single row $(s)$ is

\begin{align}
\rownll(s, t_s, r_s) &= -\log \Pr(r_s, t_s)  + \sum_{t} \pixelnll(s,t, t_s, r_s) \\
\pixelnll(s, t, t_s, r_s) &= 
\begin{cases}
 -\ln\Pr(\image(\x_{s,t})|\x_{s,t} \in R) &\text{if } t\in [t_s-r_s, t_s+r_s]\\
 -\ln\Pr(\image(\x_{s,t})|\x_{s,t}\not\in R) &\text{if } t\not\in[t_s-r_s, t_s+r_s]
\end{cases}
\end{align}

@@

\begin{align}
\ribbonnll(s, t_s, r_s) &=  \begin{cases}
-\log \Pr(r_s, t_s) & \text{if } s = 0\\
   \begin{aligned}
   & \rownll(s, t_s, r_s) \\ 
   & + \displaystyle\max_{\delta_t,\delta_r} [p(\delta_t, \delta_r)+\ribbonnll(s-1, t_s+\delta_t, r_s+\delta_r)]
   \end{aligned} & \text{if } s > 0.\\
\end{cases}
\end{align}

The quantity $p(\delta_t, \delta_r)$ is a penalty for changing direction or thickness, and could be set to the negative log likelihood of observing a change $\delta_t$ in position and a change $\delta_r$ in thickness. 
For discrete choices of $\delta_t$ and $\delta_r$ these costs can be represented by number in a matrix.

@@

% \begin{algorithmic}
% \State Initialize a matrix $A$ to same shape as ribbon image
% \For {each line of pixel $s$ in ribbon image}
%     \For{each possible pixel as $t_s$ in width of line}
%         \For{each pixel as $r_s$ in between $t_s$ and edge of line}
%             \State Store the returned probabilities from function D in $A[0, ts, rs]$
%             \If{not the first line pixel}
%                 \For {each $\delta_t and \delta_r \in \{-1, 0, 1\}$}
%                     \State Store max sum value from returned probability,
%                     \State $A[s-1, t_s+\delta_t, r_s+\delta_r]$ and penalty in $A[s, t_s, r_s]$
%                     \State Also store the value of $\delta_t$ and $\delta_r$
%                 \EndFor
%             \EndIf
%         \EndFor
%     \EndFor
% \EndFor

>>>>

@@

In order to simplify the problem, we first create a straightened ``ribbon" image in which the
sidewalk runs in a straight line, even if the original sidewalk is curved. We represent a ribbon of
length $L$ using parameters along and across to a medial curve $\mathbf{x}(s)$, where $s \in [0,
L]$. The ribbon image is the set of all points within a perpendicular distance $r(s)$ to the medial
curve.

@@

\begin{figure}[H]
    \centering \includegraphics[width=\textwidth]{Figures/sw-figure.pdf} \caption[2D Ribbon Image]{
    A demonstration of the parameters for a ribbon in 2D. $k_n$ indicates a parameter that describes
    the radius of the medial curve that we considered to be part of the ribbon, $k_f$ is a distance
    from the medial curve that we considered very unlikely to be part of the ribbon, and $k_W$ is
    the maximum width of the ribbon image.} \label{fig:2d_ribbon}
\end{figure}

As shown in figure \ref{fig:2d_ribbon}, from figure \ref{fig:osm_oxford_path}, we should assume that
there is some uncertainty in the position $\mathbf{x}_s$ of the medial curve and also uncertainty in
the radius of the ribbon. But that there is range of distances $t \in 0\dots k_n$
from the medial curve which we can accept labeling as positive examples that are within the ribbon,
and that there is some range of distances $t \in k_f\dots k_W$ which we can accept labeling as
negative sample. Finally a range $t\in (k_n+1)...(k_f-1)$ of distances that are possible locations
of the ribbon's boundary. Therefore, when searching for improved medial curves and radii of the
ribbon, we can consider a discrete range of offsets $t=-k_W,..., k_W$ chosen to be wide enough that
true ribbon image itself likely fits entirely within the grid of input negative samples along its
edges.

The orthorectified color image $\InputImage(x,y)$ records the color at the point $x,y$ in the image
plane. Then $\InputImage(\mathbf{x}_{s,t}),$ is a warped version of the image in which the ribbon is
centered on a single strip at $t=0$. The situation is illustrated in figure  \ref{fig:ribbon_3d}. In
this warped image, we can estimate the amount \OutputTrajectory{} to shift the ribbon's center in
the $t-$axis, which is orthogonal to the curve,  and also the new radius \OutputRadius{} so
that the  ribbon at each row is in columns $(\OutputTrajectory{} -\OutputRadius{}
)\dots(\OutputTrajectory{}+\OutputRadius{})$. Our problem is to estimate the most likely values of
$\OutputTrajectory{}$ and $\OutputRadius{}$.

@@

We aim to find a segmentation that combines \textit{shape} priors on the curvature and thickness of the ribbon-like features and also maximizes the likelihood of observing colors in the image. 
A variety of approaches exist to learn the conditional probability $\Pr(\Color{}|R)$ of observing a color ($\Color{}$)
 conditioned on the event ($R$) that a pixel is within a ribbon \cite{hartigan1979algorithm, jordan1999introduction}.
Many of these methods take as input a set of training examples in order to choose parameters for some model.
Our conjecture is that the initial estimate of our sidewalk's mid-line, while perhaps not perfect, is much more likely to cover pixels that are part of a sidewalk than the pixels at the edge of the ribbon image. 
We use this assumption and Bayes theorem to estimate the probability that a pixel is part of a ribbon conditioned on its color.

Let $k_n$ be a parameter that describes the radius around the medial curve that we will consider to be part of the ribbon with very high confidence. 
Let $k_f$ indicate a distance from the medial curve that we consider very unlikely to be part of the ribbon. 
Finally, let $k_W$ indicate maximum value of $t$ to consider when learning the distribution of colors in a ribbon image. 
We model the distribution of colors within the sidewalk as a Gaussian Mixture Model of $k_g=4$ full-co-variance gaussian distribution and we model the distribution of colors outside of the ribbon as a mixture $2k_g$ Gaussian since we expect more variation and complexity in this zone. 
Using the \ac{EM-GMM} algorithm, we are able to estimate $p(\Color{}|R)$ and $p(\Color{}|\neg R)$. 
Then $p(R|\Color{})=\frac{p(\Color{}|R)p(R)}{p(\Color{}|R)+p(\Color{}|\neg R)}$, where $p(R)$ is constant. 
As shown in figure \ref{fig:GMM_result}, given an initial estimate for the set of colors inside or outside of the ribbon, we estimate the weights and co-variances of the Gaussians using the \ac{EM-GMM} \cite{sridharan2014gaussian} algorithm.

@@

We will assume that widths may take on values from a set 
of predefined odd numbers (so that there is always a pixel in the center) 
Then the negative log-likelihood $\rownll(s, j, k)$ of for a 
particular offset $j$ and radius $k$ at a single row $(s)$ is

\begin{align}
\rownll(s, j, k) &= -\log \Pr(k, j)  + \sum_{t} \pixelnll(s,t, j, k) \\
\pixelnll(s, t, j, k) &= 
\begin{cases}
 -\ln\Pr(\InputImage(\x_{s,t})|R) &\text{if } t\in [j-k, j+k]\\
 -\ln\Pr(\InputImage(\x_{s,t})|\neg R) &\text{if } t\not\in[j-k, j+k]
\end{cases}
\end{align}

@@

\begin{align}
\ribbonnll(s, j, k) &=  \begin{cases}
-\ln \Pr(k, j) & \text{if } s = 0\\
   \rownll(s, j, k){+} \\
   \hspace{1em}\displaystyle\max_{\delta_t,\delta_r} 
   \left[
   -\ln p(\delta_t{,}\delta_r){+}
   \ribbonnll(s{-}1{,}j{+}\delta_t{,}k{+}\delta_r)
 \right]
 & \text{if } s > 0.\\
\end{cases}
\end{align}

The quantity $-\ln p(\delta_t, \delta_r)$ is a penalty for changing direction or thickness, and could be set to the negative log likelihood of observing a change $\delta_t$ in position and a change $\delta_r$ in thickness.

@@

% \begin{algorithmic}
% \State Initialize a matrix $A$ to same shape as ribbon image
% \For {each line of pixel $s$ in ribbon image}
%     \For{each possible pixel as $j$ in width of line}
%         \For{each pixel as $k$ in between $j$ and edge of line}
%             \State Store the returned probabilities from function D in $A[0, ts, rs]$
%             \If{not the first line pixel}
%                 \For {each $\delta_t and \delta_r \in \{-1, 0, 1\}$}
%                     \State Store max sum value from returned probability,
%                     \State $A[s-1, j+\delta_t, k+\delta_r]$ and penalty in $A[s, j, k]$
%                     \State Also store the value of $\delta_t$ and $\delta_r$
%                 \EndFor
%             \EndIf
%         \EndFor
%     \EndFor
% \EndFor
